---
title: "Course Assignment for Week3: Slides and Plotly"
author: "Mahendra S Payaal"
date: "8/15/2020"
output: ioslides_presentation
---

## Show mtcars data

### sample data shown

```{r dataShow,comments="",echo=FALSE, message = FALSE,warning=FALSE}
head(mtcars)
library(plotly)
```
## Explain Variables
### Explain variables used in plot in next slide in different stles
- wt is  weight of the car
- *mpg is miles per gallon*
- **cyl is number of cylinders**
- hp is horsepower


#### Explaination of plot 

-  Next slide will plot the weight of the car against the mileage of the car i.e. is mpg
- color by number of cylinder
- size of plots by horsepower

## Show mtcars plot 
```{r plotShow1,echo=FALSE,fig.cap = " mtcars plot",eval=TRUE}
plot(mtcars$ wt,mtcars$ mpg,xlab="weight",ylab="miles per gallon")
```


## mtcars in plotly
```{r plotShow2,echo=FALSE,fig.cap = " mtcars plot",message=FALSE,warning=FALSE}
plot_ly(mtcars, x = mtcars$wt, y = mtcars$mpg, mode = "markers", 
        color = as.factor(mtcars$cyl), size = mtcars$hp)
```

***

Thanks 

