---
title: "Capstone Project_Predict Words"
author: "Mahendra S Payaal"
date: "9/4/2020"
output: ioslides_presentation
---

## Project Outline

-  The project predicts next word(s) like in swiftkey
-  Input: one to any number of words
-  Prediction model is based on bigram, trigram and 4-gram
-  The model searches from 4-gram data set to bigrams
-  To improve accuracy upto three options are proposed
-  The project is hosted at 
   https://mspayaal.shinyapps.io/Capstone_Word_Prediction/
-  Response is instantaneous

   
## Prediction model features
- Prediction model is based on ngrams
- The search for match begins with quadgrams
- If match is found, the prediction is displayed
- If not, the search goes to trigrams and bigrams
- Thus a prediction is mostly available 
- accuracy: large database and upto 3 options  
- code at https://github.com/mspayaal/datasciencecoursera/blob/master/PredictModelCapstone.R

## Prediction code example
```{r prediction_code,echo=FALSE,warning=FALSE,message=FALSE}
library(stringr)
library(stringi)
library(dplyr)

bigram <-  readRDS( "C:/Users/mahendra.payaal/Documents/coursera/data/bigram.rds")
trigram <- readRDS("C:/Users/mahendra.payaal/Documents/coursera/data/trigram.rds")
quadgram <- readRDS("C:/Users/mahendra.payaal/Documents/coursera/data/quadgram.rds")

predictWord <- function(x){
phrase<- as.character()
  phrase <- x
  phrase <- tolower(x)
  phrase <- phrase %>%
    str_remove_all(" ?(f|ht)(tp)(s?)(://)(.*)[.|/](.*)") %>%
    str_replace_all("&amp;", "and") %>%
    str_remove_all("^RT:? ") %>%
    str_remove_all("@[[:alnum:]]+") %>%
    str_remove_all("#[[:alnum:]]+") %>%
    str_replace_all("\\\n", " ")
  phrase <- unlist(strsplit(phrase," "))
  num_words <- length(phrase)

  if (num_words == 0){
    print("No input provided")
}

  if (num_words == 1){
    phrase <- paste0("^",phrase)
    predict_1_word(phrase)
    
        }
  else if (num_words == 2){
    phrase <- paste(phrase[1],phrase[2])
    phrase <- paste0("^",phrase)
    predict_2_word(phrase)
  } 
  
  else if (num_words == 3){
    phrase <- paste(phrase[1],phrase[2],phrase[3])
    phrase <- paste0("^",phrase)
    predict_3_word(phrase)
    
  } else {
    phrase <- paste(phrase[num_words-2],phrase[num_words -1],phrase[num_words])
    phrase <- paste0("^",phrase)
    predict_3_word(phrase)
  }
  
}

predict_1_word <- function(x){
    match_list_2 <- bigram[grep(x,bigram[,1]),]
    
    if (nrow(match_list_2) == 0){ 
      print("no match")
      } 
    
      else if(nrow(match_list_2) == 1) { 
        nextword_2 <- stri_extract_last_words(match_list_2[1,1])
        print(nextword_2)
          }
      else if(nrow(match_list_2) == 2) { 
           nextword_2_1 <-      stri_extract_last_words(match_list_2[1,1])
           nextword_2_2 <- stri_extract_last_words(match_list_2[2,1])
           nextword_2 <- paste(nextword_2_1,nextword_2_2,sep=",")
           print(nextword_2)
      } else {
        nextword_2_1 <- stri_extract_last_words(match_list_2[1,1])
        nextword_2_2 <- stri_extract_last_words(match_list_2[2,1])
        nextword_2_3 <- stri_extract_last_words(match_list_2[3,1])
        nextword_2 <- paste(nextword_2_1,nextword_2_2,nextword_2_3,",")
        print(nextword_2)
          }
}

predict_2_word <- function(x){
  match_list_3 <- trigram[grep(x,trigram[,1]),]
          if(nrow(match_list_3) == 0){
          phrase_temp_1  <- unlist(strsplit(x," ")) 
          phrase_1 <- paste0("^",phrase_temp_1[2])
          predict_1_word(phrase_1)
      
      } 
      
      else if(nrow(match_list_3) == 1) { 
        nextword_3 <- stri_extract_last_words(match_list_3[1,1])
        print(nextword_3)
      }
      else if(nrow(match_list_3) == 2) { 
    nextword_3_1 <- stri_extract_last_words(match_list_3[1,1])
    nextword_3_2 <- stri_extract_last_words(match_list_3[2,1])
    nextword_3 <- paste(nextword_3_1,nextword_3_2,sep=",")
    print(nextword_3)
  } else {
    nextword_3_1 <- stri_extract_last_words(match_list_3[1,1])
    nextword_3_2 <- stri_extract_last_words(match_list_3[2,1])
    nextword_3_3 <- stri_extract_last_words(match_list_3[3,1])
    nextword_3 <- paste(nextword_3_1,nextword_3_2,nextword_3_3,sep=",")
    print(nextword_3)
  }
    }
       
predict_3_word <- function(x){
     match_list_4 <- quadgram[grep(x,quadgram[,1]),]
     
    if(nrow(match_list_4) == 0){
      phrase_temp  <- unlist(strsplit(x," ")) 
      phrase_2 <- paste(phrase_temp[2],phrase_temp[3],sep=" ")
      phrase_2  <- paste0("^",phrase_2)
      predict_2_word(phrase_2)
    
    } 
    
   else if(nrow(match_list_4) == 1) { 
      nextword_4 <- stri_extract_last_words(match_list_4[1,1])
      print(nextword_4)
    }
   else if(nrow(match_list_4) == 2) { 
     nextword_4_1 <- stri_extract_last_words(match_list_4[1,1])
     nextword_4_2 <- stri_extract_last_words(match_list_4[2,1])
     nextword_4 <- paste(nextword_4_1,nextword_4_2,sep=",")
     print(nextword_4)
   } else {
     nextword_4_1 <- stri_extract_last_words(match_list_4[1,1])
     nextword_4_2 <- stri_extract_last_words(match_list_4[2,1])
     nextword_4_3 <- stri_extract_last_words(match_list_4[3,1])
     nextword_4 <- paste(nextword_4_1,nextword_4_2,nextword_4_3,sep=",")
     print(nextword_4)
   }
}

```

- predicting on an input
- "Every inch of you is perfect from the bottom to the"
```{r prediction}
predictWord("Every inch of you is perfect from the bottom to the") 
```
- The data cleaning milestone report can be seen at 
  https://www.rpubs.com/mspayaal/653458
- 5% of all three files taken - blogs,news and tweets
- 4gram, trigram and bigram created for prediction